### 2021-12-22 Manuscript.Rmd
added ref to fig 1

---
bibliography: references.bib
csl: council-of-science-editors-author-date.csl
site: "bookdown::bookdown_site"
header-includes:
    - \usepackage{setspace}\onehalfspacing

output:
  bookdown::word_document2:
    toc: FALSE
    fig_caption: yes
---

@@

Colonial animals, consisting of genetically identical bodies that are physically and physiologically connected, can be found across the metazoan tree of life [@hiebert2020coloniality]. Functional specialization of such bodies has evolved multiple times within colonial animals, with siphonophores in particular showing the greatest diversity of functionally specialized bodies [@Beklemishev1969]. Siphonophores are highly complex, colonial “superorganisms” consisting of asexually produced bodies (termed zooids) that are homologous to solitary free-living polyps and medusae (the typical body forms in Cnidaria), but that share a common gastrovascular cavity [@dunn2006evolution;@mackie1963; @mackie1986aggregates;@mackie1987siphonophore;@totton1965synopsis]. The extreme specialization of siphonophore zooids has been of central interest to zoologists since the 19th century, in part because these zooids are highly interdependent [@Beklemishev1969; @mackie1963]. Siphonophore zooids have been likened to animal organs, with each functionally specialized zooid performing distinct roles within the colony [@mackie1963]. While this analogy works well for understanding the function of zooids within the colony as a whole, the analogy falls short in terms of explaining the evolutionary origin and development of these biological units: functionally specialized zooids are evolutionarily homologous to free living organisms, they are multicellular, possess distinct zooid specific cell types, and show regional subfunctionalization [@church2015histology; @Mackie1960; @totton1965synopsis]. While the developmental mechanisms generating zooids are very different in different clades of colonial animals [@carre1967; @carre1969etude; @carre1991complete; @dunn2006evolution; @siebert2015stem], the evolutionary processes acting on zooids may be similar to those acting on other modular biological units such as cell type, tissue, and organ [@hiebert2020coloniality]. The cellular and molecular processes underlying the patterning and molecular function of functionally specialized cnidarian zooids remain an open biological question. In recent years there has been a focus in particular on differential gene expression patterns found in different functionally specialized zooids [@10.1093/icb/icu080; @sanders2014differential; @sanders2015interspecific; @Siebert:2011gv].

@@

Using STF and SBF, combined with classical differential gene expression (DGE) analysis, we conducted DGE analyses among siphonophore zooids, and one specialized tissue (the pneumatophore, a gas filled float), in seven siphonophore species to address three questions. First, we analyzed differences in gene expression within species to identify patterns that define particular zooid types, and gain insight into the genes that may be playing a role in the structure, maintenance, and functioning of zooids. We also examined whether gene expression patterns could distinguish novel, distinct species-specific zooid types within particular species. Second, we compared gene expression between species, using STF followed by linear models to identify zooid- and species-variable genes. And third, we used phylogenetic comparative methods and SBF to identify putative clade/lineage or zooid specific expression patterns.

@@

All scripts for the analyses are available in a git repository at https://github.com/dunnlab/Siphonophore_Expression. The most recent commit at the time of the analysis presented here was fc063e9. Additional data files required for analyses that are too large for the github repository can be found on Figshare with the following DOIs: 10.6084/m9.figshare.14838384, 10.6084/m9.figshare.14838372, 10.6084/m9.figshare.14838315, 10.6084/m9.figshare.14838090, 10.6084/m9.figshare.14829183. Sequences, along with sampling metadata, are publicly available on NCBI under BioProject ID PRJNA540747.

>>>>

---
bibliography: references.bib
csl: council-of-science-editors-author-date.csl
site: "bookdown::bookdown_site"
header-includes:
    - \usepackage{setspace}\onehalfspacing
always_allow_html: true
output:
  bookdown::pdf_document2:
    keep_tex: true
    toc: FALSE
    fig_caption: yes
---

@@

Colonial animals, consisting of genetically identical bodies that are physically and physiologically connected, can be found across the metazoan tree of life [@hiebert2020coloniality]. Functional specialization of such bodies has evolved multiple times within colonial animals, with siphonophores in particular showing the greatest diversity of functionally specialized bodies [@Beklemishev1969]. Siphonophores are highly complex, colonial “superorganisms” consisting of asexually produced bodies (termed zooids) that are homologous to solitary free-living polyps and medusae (the typical body forms in Cnidaria), but that share a common gastrovascular cavity [@dunn2006evolution;@mackie1963; @mackie1986aggregates;@mackie1987siphonophore;@totton1965synopsis]. The extreme specialization of siphonophore zooids has been of central interest to zoologists since the 19th century, in part because these zooids are highly interdependent [@Beklemishev1969; @mackie1963]. Siphonophore zooids have been likened to animal organs, with each functionally specialized zooid performing distinct roles within the colony (Fig \@ref(fig:zooids-definition)) [@mackie1963]. While this analogy works well for understanding the function of zooids within the colony as a whole, the analogy falls short in terms of explaining the evolutionary origin and development of these biological units: functionally specialized zooids are evolutionarily homologous to free living organisms, they are multicellular, possess distinct zooid specific cell types, and show regional subfunctionalization [@church2015histology; @Mackie1960; @totton1965synopsis]. While the developmental mechanisms generating zooids are very different in different clades of colonial animals [@carre1967; @carre1969etude; @carre1991complete; @dunn2006evolution; @siebert2015stem], the evolutionary processes acting on zooids may be similar to those acting on other modular biological units such as cell type, tissue, and organ [@hiebert2020coloniality]. The cellular and molecular processes underlying the patterning and molecular function of functionally specialized cnidarian zooids remain an open biological question. In recent years there has been a focus in particular on differential gene expression patterns found in different functionally specialized zooids [@10.1093/icb/icu080; @sanders2014differential; @sanders2015interspecific; @Siebert:2011gv].

@@

Using STF and SBF, combined with classical differential gene expression (DGE) analysis, we conducted DGE analyses among siphonophore zooids, and one specialized tissue (the pneumatophore, a gas filled float), in seven siphonophore species to address three questions (Fig \@ref(fig:zooids-definition)). First, we analyzed differences in gene expression within species to identify patterns that define particular zooid types, and gain insight into the genes that may be playing a role in the structure, maintenance, and functioning of zooids. We also examined whether gene expression patterns could distinguish novel, distinct species-specific zooid types within particular species. Second, we compared gene expression between species, using STF followed by linear models to identify zooid- and species-variable genes. And third, we used phylogenetic comparative methods and SBF to identify putative clade/lineage or zooid specific expression patterns.

@@

All scripts for the analyses are available in a git repository at https://github.com/dunnlab/Siphonophore_Expression. The most recent commit at the time of the analysis presented here was f980b80. Additional data files required for analyses that are too large for the github repository can be found on Figshare with the following DOIs: 10.6084/m9.figshare.14838384, 10.6084/m9.figshare.14838372, 10.6084/m9.figshare.14838315, 10.6084/m9.figshare.14838090, 10.6084/m9.figshare.14829183. Sequences, along with sampling metadata, are publicly available on NCBI under BioProject ID PRJNA540747.
